5. Exponential Smoothing

### 2023-05-16 Exponential smoothing markdown draft.Rmd
finalised ETS part, changes some names

```{r cars}
#determining best ETS model
fit<-crude_oil_agg %>% 
  model(
    additive=ETS(Total_Exports~error("A")+trend("A")+season("A")),
    multiplicative = ETS(Total_Exports ~ error("M") + trend("M") + season("M")),
    random1 = ETS(Total_Exports ~ error("M") + trend("A") + season("A")),
  )

@@

@@

```{applying ETS formula}
#ETS model selection by R
crude_fit <- crude_oil_agg %>%
  model(auto_crude_ets = ETS(Total_Exports))
#checking the accuracy
report(crude_fit)
# Print the selected ETS model
print(crude_fit$auto_crude_ets)
auto_crud%>% 
  autoplot(
    crude_oil_agg) +
  labs(
    y = "Thousand Barrels",
    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))

@@

>>>>

```{r cars}
#determining best ETS model
fit<-crude_oil_agg %>% 
  model(
    additive_ETS=ETS(Total_Exports~error("A")+trend("A")+season("A")),
    multiplicative_ETS = ETS(Total_Exports ~ error("M") + trend("M") + season("M")),
    random1_ETS = ETS(Total_Exports ~ error("M") + trend("A") + season("A")),
  )

@@

addi_FC<-fit %>% select(additive) %>% forecast(h=36)

@@

```{applying ETS formula}
#ETS model selection by R
crude_fit <- crude_oil_agg %>%
  model(auto_crude_ets = ETS(Total_Exports))
#checking the accuracy and properties
accuracy(crude_fit)
report(crude_fit)
# Print the selected ETS model
print(crude_fit$auto_crude_ets)
auto_crud%>% 
  autoplot(
    crude_oil_agg) +
  labs(
    y = "Thousand Barrels",
    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))

@@

5.4 Comparison of ETS model to the simple models

```{bash}
combined_accuracy <- accuracy(crude_oil_fit) %>%
  arrange(.model) %>%
  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE) %>%
  bind_rows(
    accuracy(add),
    accuracy(crude_fit)
  )
combined_accuracy
```

Among the forecast models evaluated, the additive_ETS and auto_crude_ets models demonstrate significantly better performance compared to the Drift, Mean, Naive, and Seasonal Naive models. The additive_ETS model exhibits lower values for RMSE, MAE, MAPE, and MASE, indicating higher accuracy in predicting future values. Additionally, the model has a lower RMSSE value, suggesting a better fit to the data. While the ME is close to zero, indicating a minor bias in the forecasts, the MPE is negative, implying a slight underestimation of actual values on average. The ACF1 value of 0.169 suggests a weak positive correlation in the residuals at Lag 1. Similarly, the auto_crude_ets model shows favorable results with comparable accuracy metrics, a slightly lower ME, a higher MPE, and a slightly higher ACF1 value. Overall, the additive_ETS and auto_crude_ets models outperform the other models, making them more suitable choices for forecasting.
